% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\hypertarget{dependency-analysis-and-scheduling-exo-2-report}{%
\section{Dependency Analysis and Scheduling --- Exo 2
(Report)}\label{dependency-analysis-and-scheduling-exo-2-report}}

\textbf{Task:} In21-S8-CS4553 --- Dependency Analysis And Scheduling
(Not graded)\\
\textbf{Paper Focus:} ``Exo 2: Growing a Scheduling Language'' (Appendix
A scheduling primitives)

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-separation-of-semantics-and-scheduling-in-exo-2}{%
\subsection{1. The Separation of Semantics and Scheduling in Exo
2}\label{the-separation-of-semantics-and-scheduling-in-exo-2}}

Usually, compilers aggressively optimize source code. If a target
architecture has specific features such as vector extensions or custom
accelerators, a traditional compiler often re-organizes the source code
to utilize these extensions under the hood without changing the
semantics of the program. However, ensuring high performance often
requires intricate tuning that standard compilers struggle to safely
automate.

Exo 2 is a domain-specific scheduling language that decouples these two
aspects into distinct components: 1. \textbf{Semantics / Logic:} The
clean, reference kernel representing the fundamental algorithm. 2.
\textbf{Scheduling:} A sequence of user-defined, verified, and
semantics-preserving rewrites that target specific hardware performance
characteristics (e.g., SIMD, multithreading).

By doing this, engineers can iterate aggressively on performance without
risking the fundamental correctness of the algorithmic semantics.

Appendix A of the Exo 2 paper classifies scheduling primitives into 8
distinct categories, each designed to manipulate loops, memory, and
procedure calls safely.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{step-1-scheduling-primitive-categories-appendix-a}{%
\subsection{2. Step 1 --- Scheduling Primitive Categories (Appendix
A)}\label{step-1-scheduling-primitive-categories-appendix-a}}

\hypertarget{a.1-loop-transformations}{%
\subsubsection{A.1 Loop
Transformations}\label{a.1-loop-transformations}}

\textbf{Primitives:} \texttt{reorder\_loops}, \texttt{divide\_loop},
\texttt{divide\_with\_recompute}, \texttt{mult\_loops},
\texttt{cut\_loop}, \texttt{join\_loops}, \texttt{shift\_loop},
\texttt{fission}, \texttt{remove\_loop}, \texttt{add\_loop},
\texttt{unroll\_loop}\\
\textbf{Technique:} The goal is to reshape the overall iteration space
by safely splitting, tiling, interchanging, unrolling, separating
(fission), or merging (join) loops.\\
\textbf{Performance Gain:} This shapes the loops to map efficiently onto
specific hardware topologies. It naturally uncovers SIMD-width data
streams for the innermost loops, exposes coarse thread-level parallelism
via parallelized outer loops (pthreads), and supports complex blocking
strategies suited for GPU memory hierarchies.

\hypertarget{a.2-code-rearrangement}{%
\subsubsection{A.2 Code Rearrangement}\label{a.2-code-rearrangement}}

\textbf{Primitives:} \texttt{reorder\_stmts}, \texttt{commute\_expr}\\
\textbf{Technique:} Rearranging independent mathematical statements or
leveraging algebraic commutativity rules.\\
\textbf{Performance Gain:} This effectively mitigates artificial data
dependencies. By shifting instructions, compilers can form denser
patterns suitable for instruction replacement or advanced vectorization
schemas.

\hypertarget{a.3-scope-transformations}{%
\subsubsection{A.3 Scope
Transformations}\label{a.3-scope-transformations}}

\textbf{Primitives:} \texttt{specialize}, \texttt{fuse},
\texttt{lift\_scope}\\
\textbf{Technique:} Restructuring or relocating entire control scopes
safely.\\
\textbf{Performance Gain:} Scope fusion significantly increases cache
locality. Specialization eradicates redundant conditional checks by
resolving overhead at compile time when specific constraints are
mathematically proven. Scope lifting resolves complex nesting structures
introduced post-tiling operations.

\hypertarget{a.4-multiple-procedures}{%
\subsubsection{A.4 Multiple Procedures}\label{a.4-multiple-procedures}}

\textbf{Primitives:} \texttt{inline}, \texttt{replace},
\texttt{call\_eqv}, \texttt{extract\_subproc}\\
\textbf{Technique:} Executing inter-procedural optimizations, custom
instruction selection mappings, or isolating specific algorithm subsets
into sub-kernels.\\
\textbf{Performance Gain:} The key component of micro-kernel design,
this allows developers to systematically replace slow scalar loops with
heavily optimized hardware vector intrinsics (such as AVX2).
Furthermore, it allows substituting equivalent routines while ensuring
complete semantic parity.

\hypertarget{a.5-buffer-transformations}{%
\subsubsection{A.5 Buffer
Transformations}\label{a.5-buffer-transformations}}

\textbf{Primitives:} \texttt{lift\_alloc}, \texttt{sink\_alloc},
\texttt{delete\_buffer}, \texttt{reuse\_buffer}, \texttt{resize\_dim},
\texttt{expand\_dim}, \texttt{rearrange\_dim}, \texttt{divide\_dim},
\texttt{mult\_dim}, \texttt{unroll\_buffer}, \texttt{bind\_expr},
\texttt{stage\_mem}\\
\textbf{Technique:} Modifying buffer allocation locations, altering
underlying dimensionality, shapes, or data layouts, and staging memory
windows directly into hardware temporary registers.\\
\textbf{Performance Gain:} Critical for enforcing strict cache and
shared-memory tiling behaviors (via \texttt{stage\_mem}). It ensures
arrays fit perfectly within constrained register spaces (\texttt{expand}
or \texttt{divide}). Decreases the general application memory footprint
(\texttt{sink} combined with \texttt{resize}) and drastically improves
efficiency by guaranteeing the reuse of buffers across distinct
operational lifetimes.

\hypertarget{a.6-simplification}{%
\subsubsection{A.6 Simplification}\label{a.6-simplification}}

\textbf{Primitives:} \texttt{simplify}, \texttt{eliminate\_dead\_code},
\texttt{rewrite\_expr}, \texttt{merge\_writes}, \texttt{inline\_window},
\texttt{inline\_assign}\\
\textbf{Technique:} Streamlining loops to clean up excessive overhead
generated strictly from the preceding transformation steps.\\
\textbf{Performance Gain:} Results in tighter, streamlined generated
code. It directly eradicates structurally redundant loads, duplicate
writes, or dead-code paths.

\hypertarget{a.7-backend-checked-annotations}{%
\subsubsection{A.7 Backend-Checked
Annotations}\label{a.7-backend-checked-annotations}}

\textbf{Primitives:} \texttt{set\_memory}, \texttt{set\_precision},
\texttt{parallelize\_loop}, \texttt{set\_window}\\
\textbf{Technique:} Explicitly adding precise data annotations that the
backend compiler code generator natively verifies.\\
\textbf{Performance Gain:} It actively guarantees that loops selected
for multi-threading are strictly memory-safe for parallel execution. The
annotations are also essential for forcing specific SIMD intrinsics,
correctly aligning precision types, and guaranteeing strict hardware ABI
correctness.

\hypertarget{a.8-configuration-state}{%
\subsubsection{A.8 Configuration State}\label{a.8-configuration-state}}

\textbf{Primitives:} \texttt{bind\_config}, \texttt{delete\_config},
\texttt{write\_config}\\
\textbf{Technique:} Directly scheduling the underlying configuration
state writes strictly geared toward stateful hardware accelerators.\\
\textbf{Performance Gain:} Heavily reduces structural state overhead by
successfully hoisting out generic configuration setup phases and
eliminating duplicated or redundant state writes occurring entirely
within the innermost hardware loop limits.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{step-2-install-exo-and-run-examples}{%
\subsection{3. Step 2 --- Install Exo and run
examples}\label{step-2-install-exo-and-run-examples}}

\hypertarget{install-pip}{%
\subsubsection{Install (pip)}\label{install-pip}}

pip install exo-lang

\hypertarget{or-build-from-source}{%
\subsubsection{Or build from source}\label{or-build-from-source}}

git clone https://github.com/exo-lang/exo.git cd exo git submodule
update --init --recursive python -m venv \textasciitilde/.venv/exo
source \textasciitilde/.venv/exo/bin/activate python -m pip install -U
pip setuptools wheel python -m pip install -r requirements.txt python -m
build . pip install dist/*.whl

Run quizzes: - exo/examples/quiz1: exocc quiz1.py - exo/examples/quiz2:
exocc quiz2.py - exo/examples/quiz3: exocc quiz3.py

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{step-3-quiz-scenarios-and-solutions}{%
\subsection{4. Step 3 --- Quiz Scenarios and
Solutions}\label{step-3-quiz-scenarios-and-solutions}}

The following are descriptions of the issues encountered in the provided
scheduling quizzes, along with analyses of the incorrect approaches and
the verified, correct solutions.

\hypertarget{quiz-1-avx2-replacement-not-occurring}{%
\subsubsection{Quiz 1 --- AVX2 Replacement Not
Occurring}\label{quiz-1-avx2-replacement-not-occurring}}

\textbf{Issue:} The generated schedule correctly mimics the structure of
SIMD vector code but fails to invoke the underlying hardware vector
intrinsics. It is still executing scalar operations one element at a
time. \textbf{Incorrect Approach:} Invoking the
\texttt{replace\_all(p,\ avx\_instrs)} schedule directive \emph{before}
correctly defining the memory topology of the target staged buffers.
\textbf{Correct Solution:} The \texttt{replace\_all} function in Exo is
memory-aware. Therefore, you must invoke
\texttt{set\_memory(p,\ f"\{name\}\_vec",\ AVX2)} on all intermediate
buffer arrays before performing the replacement. Once the memory context
is explicitly \texttt{AVX2}, the compiler pattern-matches and replaces
the statements with the correct \texttt{\_mm256\_*} intrinsic mappings.
\textbf{Source:} \texttt{solutions/quiz1\_solution.py}

\hypertarget{quiz-2-schedulingerror-during-loop-fission}{%
\subsubsection{Quiz 2 --- SchedulingError During Loop
Fission}\label{quiz-2-schedulingerror-during-loop-fission}}

\textbf{Issue:} The compiler throws a \texttt{SchedulingError} stating
``Will not fission here, because doing so will hide the allocation of
vec from a later use site.'' \textbf{Incorrect Approach:} Utilizing a
single loop iteration to immediately perform \texttt{expand\_dim},
\texttt{lift\_alloc}, and \texttt{fission} sequentially for each
temporary array. This hides the earlier allocations from subsequent uses
because the memory allocations remain trapped within the split loop
bodies. \textbf{Correct Solution:} Refactoring the scheduling code to
operate in a 2-pass schema. First, loop over the structures and run
\texttt{expand\_dim} and \texttt{lift\_alloc} on all temporary
allocations to safely hoist them outside the computation context.
Secondly, safely run a subsequent loop to \texttt{fission} the
individual assignment operations. \textbf{Source:}
\texttt{solutions/quiz2\_solution.py}

\hypertarget{quiz-3-blur_x-width-fails-to-shrink}{%
\subsubsection{\texorpdfstring{Quiz 3 --- \texttt{blur\_x} Width Fails
to
Shrink}{Quiz 3 --- blur\_x Width Fails to Shrink}}\label{quiz-3-blur_x-width-fails-to-shrink}}

\textbf{Issue:} While attempting to tile a blur operation, the scheduler
successfully shrinks the overall allocation height dimension to 34 but
entirely fails to dynamically reduce the width mapping down to 256.
\textbf{Incorrect Approach:} The helper function
\texttt{get\_loops\_at\_or\_above(cursor)} initializes its traversal
collection as \texttt{loops\ =\ {[}{]}}. This fundamentally omits the
initial underlying cursor (the innermost \texttt{xo} loop).
Consequently, the intermediate array allocation is never sunk deep
enough into the \texttt{xo} scope. \textbf{Correct Solution:} Modifying
the initialization definition to \texttt{loops\ =\ {[}cursor{]}}
correctly captures the \texttt{xo} loop mapping inside the traversal
collection. This allows the allocation to gracefully sink into the
correct inner tile representation, reducing the active layout
requirements perfectly down to \texttt{{[}34,\ 256{]}}. \textbf{Source:}
\texttt{solutions/quiz3\_solution.py}

\end{document}
